Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-15 manuscript.tex
Update on Overleaf.

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Why do neurons communicate through spikes? By definition, spikes are all-or-none neural events which occur at continuous times. In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock. This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks. As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data. With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming. In this review paper, we provide evidence from biology and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Is there a neural code?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CONTEXT
Neural activity is directly influenced by our immediate environment and by internal states, and is structured in order to generate motor actions. The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection. A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings that can be observed in neurobiological experiments. One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e. an explicit representation within the neural activity. Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity. In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}). A central and well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}. In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons. First, these have the particular property of being essentially binary in their amplitude, i.e. they are prototypical, "\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}". Typically, these are received by a neuron from other afferent neurons on the arborized "input" dendrite, integrated in its soma, and possibly further propagated along its "output" axon to reach efferent neurons The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code. Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons. This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.

@@

However, it was not until the study of the interaural time difference (ITD) by~\citet{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function. This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears. According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris. A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e. the temporal disparity between left and right cochlear loci in the owl's brain. However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears. The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a "coarse" regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection. These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.

@@

\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e. the amount of myelin wrapped around the axon. Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}. Myelinization consists in "two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions"~\citep{simons_oligodendrocytes_2016}. Multiple myelin regions can appear on one neuron axon and form the following subdomains: "the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)"~\citep{duncan_2021}. In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}. This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}. Thus, in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. axonal delay regulation.

@@

It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}. Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}. ~\citet{duncan_neuron-oligodendrocyte_2021} states that "the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration".

@@

% myelination for axonal delay, but what about dendritic delay ? 
The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination. However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}. Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by "both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output". The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.

@@

An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process. Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. In that work, based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation. They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing. While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Learning synaptic delays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Training delays ~\citep{state_training_2019}
First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Real-world applications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios. Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility. For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream. The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern. The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.

@@

To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}. It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate. Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, like it can be observed in biological neurons. Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons. The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector". Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms. For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones. SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations. Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning. This flexibility in the implementation comes at the cost of an increased energy consumption. Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}. They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction. BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}. It is a wafer-scale neuromorphic hardware with analog components. Analog arrays (i.e. field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices. The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems. Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.

@@

Online on-chip learning and computations with delays are still emerging in neuromorphic engineering. The technical challenges linked to the development of this type of implementations and the growing interest for delay learning makes advances in this field interesting for the future of computations with precise spatio-temporal motifs. While improvements are still to be made, neuromorphic chips seem to be good candidate to efficiently make use of these particular features.

%they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing. In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible

%review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)

%We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method. Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022}

%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160

% solutions usuellement utilisées - surrogate gradient
%for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Why do neurons communicate through spikes? By definition, spikes are all-or-none neural events which occur at continuous times. In other words, spikes are on one side binary, existing or not without further details, and on the other can occur at any asynchronous time, without the need for a centralized clock. This stands in stark contrast to the analog representation of values and the discretized timing classically used in digital processing and at the base of modern-day neural networks. As neural systems almost systematically use this so-called event-based representation in the living world, a better understanding of this phenomenon remains a fundamental challenge in neurobiology in order to better interpret the profusion of recorded data. With the growing need for intelligent embedded systems, it also emerges as a new computing paradigm to enable the efficient operation of a new class of sensors and event-based computers, called neuromorphic, which could enable significant gains in computation time and energy consumption --- a major societal issue in the era of the digital economy and global warming. In this review paper, we provide evidence from biology, theory and engineering that the precise timing of spikes plays a crucial role in our understanding of the efficiency of neural networks.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction: Importance of precise spike timings in the brain}\label{sec:time}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Is there a neural code?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CONTEXT
Neural activity is directly influenced by our immediate environment and by internal states, and is structured in order to generate motor actions. The efficiency of these actions is key for survival, which is the sole determinant in the light of natural selection. A central question of modern neuroscience is to better understand the essence of neural activity, as exemplified by the recordings that can be observed in neurobiological experiments. One sometimes uses the expression ``decoding the neural code'', although this implies the existence of a code, i.e. an explicit representation within the neural activity. Nevertheless, we will use this terminology in all generality to denote the existence of a structure in neural activity. In this respect, it is known that neural activity may be related to specific measurable variables, and since Galvani's experiments, we know that an electrical activity applied on muscular nerves can cause the stretching of a frog's limb (for a review, see~\citep{piccolino_luigi_1997}). A central and well studied way of communication between neurons are specific electrochemical events called action potentials, or spikes, which were first discovered by \href{https://en.wikipedia.org/wiki/Edgar_Adrian}{Lord Adrian}~\citep{adrian_impulses_1926}. In the scope of this article, we focus mainly on these spikes, which are brief (about one millisecond) impulses that propagate along the axons of neurons. First, these have the particular property of being essentially binary in their amplitude, i.e. they are prototypical, 
``\href{https://en.wikipedia.org/wiki/All-or-none_law}{all-or-none}''. Typically, these are received by a neuron from other afferent neurons on the arborized ``input'' dendrite, integrated in its soma, and possibly further propagated along its ``output'' axon to reach efferent neurons. The frequency at which these spikes are emitted has been shown to be roughly commensurate with the stretch of the frog's limb~\citep{adrian_impulses_1926}, and this variable (measured in spikes per second, or Hertz) is as of today considered to be the main constituent of the neural code. Indeed, neurophysiologists typically use the temporal evolution of the firing rate (for instance, as computed as the average occurrence of spikes in small temporal windows of about $100~\ms$) in order to characterize the dynamical activity of neurons. This may be extended by computing different statistics on each neuron's sequence of spikes~\citep{perkel_neuronal_1967} but also the dependence across neurons~\citep{perkel_neuronal_1967-1}.

@@

However, it was not until the study of the interaural time difference (ITD) by~\citet{gerstner_neuronal_1996} that it was discovered that this delay is not homogeneous for all neurons of the same type and species, but adapts according to their function. This ITD is a biological mechanism which allows for the azimuthal localization of sound by barn owls, by organically computing the difference in arrival time of a sound between their two ears. According to~\citep{gerstner_neuronal_1996}, there is a true paradox in auditory neural systems since “neural networks encode behaviorally relevant signals in the range of a few $\mu s$ with neurons that are at least one order of magnitude slower”, therefore necessarily need to play on the response time to do so.~\citet{pena_cochlear_2001} conducted an extended study on ITD in order to confirm this theory by identifying the coincidence detector involved, which the authors situated in the nuclear laminaris. A first hypothesis suggested that the sound coincidence was detected using stereausis, i.e. the temporal disparity between left and right cochlear loci in the owl's brain. However, it was quickly set aside as the predictions did not match the measured disparities in the loci, and no variation was perceived in the nuclear laminaris for a similar sound intensity in both owl ears. The authors support a second hypothesis, that of different axonal delays in the ipsi- and contra-lateral cochlear nucleus magnocellularis.~\citet{seidl_mechanisms_2010} experimentally seconds this hypothesis of a ``coarse'' regulation of delay, as the authors conclude that regulations at different sites within individual axons of at least two parameters, namely the axon diameter and internode distances, might be responsible for precise adjustments of physiological delays thus allowing the ITD detection. These experiments also confirmed previous studies stating that the conduction velocity of a spike in a neuron (in other words, its delay) depends strongly on the axon diameter~\citep{gasser_axon_1939} and the internode distance between Ranvier nodes~\citep{brill_conduction_1977} --- instead of solely the axon's length as was previously thought.

@@

\citet{gasser_axon_1939} experimentally confirmed with homogeneously selected neurons that the axonal delay is positively proportional to the axon diameter, i.e. the amount of myelin wrapped around the axon. Indeed, the oligodendrocytes, one of the many glial cells present in the Vertebrates' nervous system identified in 1924 by Pío del Rigo Hortega~\citep{perez-cerda_pio_2015}, produce thin protein sheets interspersed with lipid layers wrapped concentrically around the axon, called myelin~\citep{schmitt_ultrastructure_1939}. Myelinization consists in ``two motions: the wrapping of the leading edge at the inner tongue around the axon underneath the previously deposited membrane and the lateral extension of myelin membrane layers toward the nodal regions''~\citep{simons_oligodendrocytes_2016}. Multiple myelin regions can appear on one neuron axon and form the following subdomains: ``the internode (corresponding to the compacted region of myelin), the paranodes (where the outer loops of the myelin contact the axon), the juxtaparanode (the interface between the paranode and compact myelin, rich in potassium channels) and the node of Ranvier (the $\backsim$1 $\mu m$ gap between adjacent myelin internodes [allowing for] the saltatory conduction)''~\citep{duncan_2021}. In average, each oligodendrocyte produces 20 to 60 myelinization processes and each myelin sheath is 20 to 200 $\mu m$ long~\citep{simons_oligodendrocytes_2016}. This demonstrates an additional impact of myelinization on the conduction velocity, as the number of segments is positively correlated to the axonal delay~\citep{brill_conduction_1977}. Thus, in several occasions, myelin has been identified as a main actor in the regulation of conduction velocity in neurons, i.e. axonal delay regulation.

@@

It is worth highlighting that myelinization becomes increasingly important in larger brains where conductance delays are substantial and brainwave rhythms are critical; synchrony errors can lead to neuropsychiatric and neurological dysfunctions~\citep{fields_new_2015}, such as Parkinson's disease, epilepsy or multiple sclerosis~\citep{madadi_asl_dendritic_2018}. Additionally, a recent study suggests that demyelination of the optic nerve could be an underlying factor in glaucoma~\citep{xue_demyelination_2021}. ~\citet{duncan_neuron-oligodendrocyte_2021} states that ``the loss of myelin and oligodendrocytes fundamentally alters the neuron, [which are then] susceptible to energetic failure [and] subsequent degeneration''.

@@

% myelination for axonal delay, but what about dendritic delay ? 
The previous paragraphs present the biological mechanisms behind axonal delay, regulated by myelination. However, dendritic delay seems to play an equally important role in the precise timing within a sequence of spikes: dendritic delay is involved in the performance, structure and function of the nervous system, the modulation of spatio-temporal properties of pre- and post-synaptic activity patterns and the functional limitations of sensory feedback control efficiency~\citep{madadi_asl_dendritic_2018}. Its role has been specifically identified in the compensation of input asynchrony in the mammalian auditory brainstem~\citep{spencer_compensation_2018}.~\citet{branco_dendritic_2010} demonstrates the dendritic sensitivity to a sequence of synaptic activation in cortical pyramid neurons, encoded by ``both local dendritic calcium signals and somatic depolarization, leading to sequence-selective spike output''. The dendritic mechanism described can identify patterns delivered to a single dendrite or randomly distributed across the dendritic tree, and relies on the dendritic calcium influx moderation by NMDA receptors.

@@

An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process. Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. In that work, based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation. They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing. While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
\subsection{Learning synaptic delays}
% Training delays ~\citep{state_training_2019}
First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Real-world applications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%----------------------------%
%
\begin{figure}%[t!]
  \centering
  \includegraphics[width=0.980\linewidth]{figures/izhikevitch_repro.png}
    \caption{Detecting motion in an event stream with heterogeneous delays. 
     }
  \label{fig:model}
\end{figure}
% 
%----------------------------%

A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios. Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility. For instance, in a recent work,~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream. The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern. The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.

@@

To our knowledge, the first neuromorphic circuit is the pulsed current-source synapse proposed by Carver Mead in 1989~\citep{mead_analog_1989}. It was implemented with transistors operating in the sub-threshold domain and responded to asynchronous events, but was not capable of discriminating two different spiking sequences with the same firing rate. Indeed, the postsynaptic membrane potential was increased by a step proportional to the input current but did not decrease in time, like it can be observed in biological neurons. Then, electronics circuits became more and more bio-realistic and, two decades later, ~\citep{bartolozzi_synaptic_2007} released the Diff-Pair Integrator (DPI) synapse that could reproduce the global dynamics of the biological neurons. The DPI circuit could multiplex in time spikes from different sources and became a potential  ``silicon coincidence detector''. Today many devices are good candidates for implementing event-based algorithms and use the AER, they can be divided intro 3 major categories: digital, analog and mixed analog/digital platforms. For a more complete review the reader can refer to~\citep{schuman_survey_2017}, here we site the most popular ones. SpiNNaker~\citep{furber_overview_2013, furber_spinnaker_2020}, Loihi~\citep{davies_loihi_2018} and TrueNorth~\citep{merolla_million_2014} chips are widely used examples of digital hardware implementations. Compared to TrueNorth which exclusively implements a LIF neuron, SpiNNaker and Loihi offer some flexibility in terms of neuron model and allow for on-chip learning. This flexibility in the implementation comes at the cost of an increased energy consumption. Mixed-analog-digital systems were developed at the Stanford University: Neurogrid and Braindrop~\citep{benjamin_neurogrid_2014, neckar_braindrop_2019}. They are mostly used by computational neuroscientist to model brain activity with different levels of abstraction. BrainScaleS~\citep{schemmel_wafer-scale_2010} is another mixed-analog-digital system developed, just like SpiNNaker, for the Human Brain Project~\citep{markram_introducing_2011}. It is a wafer-scale neuromorphic hardware with analog components. Analog arrays (i.e. field programmable analog arrays (FPAA)) refer to the initial idea of neuromorphic hardware aiming at building strictly analog devices. The pulsed current source synapse and the DPI are examples of such devices, we can also mention the field programmable neural array~\citep{farquhar_field_2006} and the NeuroFPAA~\citep{cheng_fpaa_2009} specifically designed for neuromorphic systems. Due to their lack of generality and some issues specific to analog circuits, these fully analog devices are not yet widely used for neuromorphic computing.

@@

Online on-chip learning and computations with delays are still emerging in neuromorphic engineering. The technical challenges linked to the development of this type of implementations and the growing interest for delay learning makes advances in this field interesting for the future of computations with precise spatio-temporal motifs. While improvements are still to be made, neuromorphic chips seem to be good candidate to efficiently make use of these particular features. 
%
%they made a Spike-Timing Neuromorphic Architecture with a dendritic arbor to implement the delays and a transfer function for the weights which is well-suited for areas such as pattern recognition and natural language processing. In this paper, we formally describe the STPU, implement the STPU on a field programmable gate array, and show measured performance data.\citep{hill_spike-timing_2017} not accessible
%
%review on neuromorphic computations for robotic\citep{sandamirskaya_neuromorphic_2022} (not accessible for the moment)
%
%We apply delay-based reservoir computing as the information processing core, along with a novel training and labelling method. Different from the conventional ECG classification techniques, this computation model is a end-to-end dynamic system that mimics the real-time signal flow in neuromorphic hardware~\citep{liang_neuromorphic_2022} 
%
%Real-time cortical simulation on neuromorphic hardware -> https://royalsocietypublishing.org/doi/full/10.1098/rsta.2019.0160
%
% solutions usuellement utilisées - surrogate gradient
%for other section: A hypothesis for temporal coding of young and mature granule cells(https://www.frontiersin.org/articles/10.3389/fnins.2013.00075/full)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%: >>>>>>>> Laurent is here <<<<<<<<<<<<<<
In this review paper, we have presented recent evidence for the role of precise spiking motifs in neuroscience. In particular, we have shown that such particular motifs may play a crucial role in neurobiology, that they may be understood at the theoretical and computational levels and that they may have numerous applications in neuromorphic engineering. In particular, we have shown that:
\begin{itemize}
  \item  the efficiency of neural systems, and in particular the visual system, imposes strong constraints on the structure of neural activity which highlight the importance of precise spike times,
  \item  growing evidence from neurobiology proves that neural systems are more than integrators and may use synchrony detection in different forms: synfire chains, traveling waves on arbitrary spiking motifs,
  \item  many theoretical models already exist taking into account the specificity of spiking motifs, notably by using heterogeneous delays,
  \item  ultimately that using precise puking motifs could be a key ingredient in neuromorphic systems to reach similar efficiencies as biological neural systems.
\end{itemize}
 Overall, our reviewing effort has shown that a growing community is focusing on that aspect. This community is based on solid and validated evidence which is getting novel grounds thanks to current technical advances. Moreover, we have also shown that this community is highly diverse, operating on \emph{in vivo}, \emph{in computo} and \emph{in silico} systems. As a consequence, the global effort is still largely scattered, which limits its larger acceptance in neuroscience.
